General linear Models Coursework

### 2022-11-20 Draft_final.Rmd
Updating inference on sex and household size (simplifing as now have table)

## Part 4: Regressing the natural logarithm of expenditure on income
Fitting the same model as in task 2 but with the response of expenditure logged (a log-linear model) $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the below estimated coefficients and standard errors.

@@

@@

The final 2 variables, sex of household head and household size have an interaction between them. However, only the household size of 3 level has a statistically significant interaction with sex of household head. Therefore as there is no evidence the other levels of the interaction are different from the reported associations without the interaction included I will not report them.

The sex of the household head is highly statistically significant with being female head having a 95% CI of [1.056, 1.259] which corresponds to a **5.6%** to **25.9%**  increase the expenditure compared to a male head with all other variables held constant (and for a household not of size 3). Household size is also extremely statistically significant with a 95% CI of [1.224, 1.455], [1.313, 1.668] and [1.349, 1.816] for households of size 2, 4 and 5+ respectively when compared to a household of size 1 with all other variables held constant which corresponds to a **22.4%** to **45.5%**, a **31.3%** to **66.8%** and a **34.9%** to **81.6%** respectively.

#creating table of 95% CI calculated
data.frame(t(data.frame("Male"=male.CI, "Female"=female.CI)))%>%
  rename("1 Person"="X1",
         "2 Persons"="X2",
         "3 Persons"="X3",
         "4 Persons"="X4",
         "5+ Persons"="X5")%>%
  kable(caption = "95% CI compared to male headed household of 1 person",format = 'pandoc')
```

```{r,include=F}
# #creating table of 95% CI calculated
# kable(data.frame(data.sex.household.CI), caption = "95% CI compared to male headed household of 1 person")%>%
#   row_spec(0,bold=T)%>%
#   column_spec(1,bold=T)%>%
#   kable_classic(position = "center",latex_options = "HOLD_position")
# 
# knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=3)%>%
#   row_spec(0,bold=T)%>%
#   kable_styling(position = "center",latex_options = "HOLD_position")
```

For the effect on expenditure between a household of size 1 and size 3 because of the significant interaction with sex of household head if the household head is male the 95% CI is [1.362, 1.719] and [0.912, 1.920] if female with all other variables held constant. Note that for the female head 1 is included in the CI, therefore there is no evidence of a difference in expenditure between a household of size 1 with a female head and a household of size 3 with a female head with all other variables held constant. However for a for a male household head, a household of size 3 compared to size 1 corresponds to a **36.2%** to **71.9%** increase in expenditure with all other variables held constant.

@@

Further investigation may be required for tenure type as it could be other variables which cause a difference in tenure are responsible.

>>>>

## Part 4: Regressing the natural logarithm of expenditure on income
Fitting the same model as in task 2 but with the log of the response expenditure (a log-linear model) $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the below estimated coefficients and standard errors.

@@

Sex of household head and household size have an interaction between them therefore cannot be reported separately. To easily display the effects of both on expenditure I created the below table which displays the multiplicative effect of sex and household size on expenditure.

#creating table of 95% CI calculated
data.frame(t(data.frame("Male head"=male.CI, "Female head"=female.CI)))%>%
  rename("1 Person"="X1",
         "2 Persons"="X2",
         "3 Persons"="X3",
         "4 Persons"="X4",
         "5+ Persons"="X5")%>%
  kable(caption = "95% CI of multiplcative effect on expenditure compared to male headed household of 1 person",format = 'pandoc')

table(expend.df$sex.hh,expend.df$hh.size)
```

For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1 however, for a household of size 3 or 5+ there is not enough evidence to support a difference in expenditure (CI includes 1). I note the confidence intervals are very large thus there is alot of uncertainty in these results and thus the reason for no found effect may be due to insufficient power of thr sample.

Compared to a household of size 1 all larger households increase expenditure if household head is male whereas due to the uncertainty there is no evidence this is the case for female household heads due to overlapping CI.

@@

@@

Further investigation is required for the affect of sex of household head as due to the small sub samples t

may be required for tenure type as it could be other variables which cause a difference in tenure are responsible.
